\sysfull

### 2018-08-06 paper/asplos19/background.tex
backgroud pragraph cut on FRAM

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{WAR-affected code execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state. \note{replace checkp(); with checkpoint();} \note{reduce font size by one point}}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}. Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}. Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory. Fig.~\ref{fig:war} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation. If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9). \note{in the above example it should \ref{fig:war-example}, Line 6, first and then \ref{fig:war-execution}, Lines 5 and 8}

@@

\textbf{Direct Non-volatile Memory Access.}  \note{I want to emphasize again that we do not evaluate NV memory accesses} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency. Checkpointing systems copy all volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at every checkpoint.  Task-based models, which do not take checkpoints, back up a subset of non-volatile data to maintain memory consistency using, e.g., redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.  Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile, perform high-frequency non-volatile memory accesses, checkpointing frequently. On the other hand, \sys accesses non-volatile memory by using only DMA and without CPU intervention: \sys moves a batch of variables from non-volatile memory into volatile memory and programs are only allowed to access individual variables in volatile memory directly.

>>>>

\begin{figure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
        \caption{WAR-affected code}
        \label{fig:war-example}
    \end{subfigure}
    \begin{subfigure}{0.49\columnwidth}
        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
        \caption{WAR-affected code execution}
        \label{fig:war-execution}
    \end{subfigure}
    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state.}
    \label{fig:war}
\end{figure}

@@

A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}. Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}. Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory. Fig.~\ref{fig:war} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation. If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 6), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 5 and 8).

@@

%\textbf{Direct Non-volatile Memory Access.} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency. Checkpointing systems copy all volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at every checkpoint.  Task-based models, which do not take checkpoints, back up a subset of non-volatile data to maintain memory consistency using, e.g., redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.  Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile, perform high-frequency non-volatile memory accesses, checkpointing frequently. On the other hand, \sys accesses non-volatile memory by using only DMA and without CPU intervention: \sys moves a batch of variables from non-volatile memory into volatile memory and programs are only allowed to access individual variables in volatile memory directly.
